% Please do not change the document class
\documentclass{scrartcl}

% Please do not change these packages
\usepackage[hidelinks]{hyperref}
\usepackage[none]{hyphenat}
\usepackage{setspace}
\doublespace

% You may add additional packages here
\usepackage{amsmath}
\usepackage{graphicx} 
\graphicspath{ {Diagrams/} }

% Please include a clear, concise, and descriptive title
\title{Dissertation Notes }

% Please do not change the subtitle
%\subtitle{COMP240}

% Please put your student number in the author field
\author{1507866}

\begin{document}
	
\maketitle

 \section{Ideas} 
  Hollywood AI: Visualising AI decision making as part of the game play experience. Possible questions:
 \begin{itemize}  
 	\item 
 \end{itemize}

\section{Assignment Brief Questions}
 \begin{itemize}  
	\item What is (are) the key research question(s) that you will seek to answer in
	your project?
	\item How will answering these questions contribute to the state of knowledge
	in the field of your project?
	\item What methodology will you use to seek answers to these questions?
	Justify your methodology.
	\item What preliminary results have you obtained?
\end{itemize}

\section{Games:}
 \begin{itemize}  
	\item Third Eye Crime
	\item Shadwen
\end{itemize}


 \section{Papers:}
 \subsection{AI-based game design patterns  \cite{treanor2015ai} }
 
 \begin{itemize}  
 	\item  About designing games are AI not designing AI to fit a game?
 	\item Game AI normally rudimentary because complexity not required 
 	\item `` We define foreground  AI as  agents  the  player  notices  and  can  reason  about.''
 	\item 	Ai is visualized - based around manipulating AI state 
 	\item
 \end{itemize}

 \subsection{Visualizing decision trees in games to support children's analytic reasoning: any negative effects on gameplay?  \cite{Haworth2010} }

 
\bibliographystyle{ieeetr}
\bibliography{comp320}

	
\end{document}
